Universal Anonymous Signatures

### 2022-05-10 prism-backend/research/gsac/introduction.tex
Added sketch of security proof for SBCM. Updated GSAC model to include attrs on Open. Other minor edits.
Working on intro.

\commentwho{Jesus}{The following will most probably change completely if we
  are to submit this for publication. However, this should capture tha main
  ideas.}

In this work, we combine two closely related schemes -- group signatures (GS)
and anonymous credentials (AC) -- which have followed a
separate path up to now, beyond individual constructions from both fields
drawing inspiration (and techniques) from one another. This combination is in
some sense a natural evolution, which we do believe may have relevant impact in
practical settings. On the one hand, anonymous credentials enable a richer
functionality in regards to what one can prove about his own identity, in a
privacy-preserving way. Academic proof of this is the fact that AC systems have
naturally evolved to enabling arbitrary claims that only reveal the minimally
possible 1-bit of information \needcite. Also, AC systems have been proposed to
go even beyond what one can prove about his identity, covering also orthogonal
functionality such as delegation \needcite, or aggregation of multiple
credentials into one \needcite. Even outside academia, prominent
technical workgroups \needcite, startups \needcite, and big techs \needcite are
promoting and adopting solutions which are directly inspired by AC systems. On
the other hand, GS 
schemes have inspection functionality (e.g., opening and linking of signatures)
at their core, which is reflected formally through non-frameability properties.
This has originated a large body of research focusing on the construction and
formalisation of variants of these notions \needcite. Consequently, group
signatures tend to be proposed for scenarios that require some sort of
de-anonymization \needcite. Indeed, this is not possible in ACs, which at most
support revocation of issued credentials, which essentially can compare to
linking, but not to opening-related notions.

Our goal is to get the best of both worlds: leverage the richer functionality
of ACs in what respects to proving, in a privacy-preserving manner, useful
claims on one's identity; and, at the same time support, through inspection
capabilities, use cases that may require some sort of de-anonymization. This is
not the first work that somehow follows this direction, though --
Attribute-Based Group Signatures, covered in \secref{ssec:related}, propose
group signatures with attributes (although mentions to AC works from the ABGS
papers are non-existent, so the resemblance seems circumstancial). However, we
believe that ABGS lack the flexibility that has made AC schemes gain popularity
recently. Hence, we propose a different approach that directly integrates GS
features and properties with those of AC, as closely as possible to their
original notions.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "sok-privsig"
%%% End:

>>>>

% What is the context for this research? Introduce previous works on this topic,
% in one paragraph. (Privacy-preserving identities, with accountability (APPI))
Successful identity representation, management, and usage are, on its own, among
the toughest problems in the Internet and digital communications at large
\needcite. Just as an illustration, think
of usernames and passwords, which end up being unmanageable as soon as we need
to start remembering a few \needcite; or traditional vertical PKIs, like X.509,
which are too rigid, and frequently require too much trust \needcite; or
horizontal PKIs, such as PGP, that are also too rigid, and do not scale well
\needcite. If, in addition, we require privacy, the scenario becomes even more
challenging. Technologies like OpenID and OAuth \needcite make identity more
manageable (from the point of view of service providers) through Single Sign
On solutions. But this comes at the cost of introducing centralization points
that gain way too much knowledge beyond what is actually necessary. Holding
users accountable, or extracting utility out of their actions, is trivial in
this cases -- although that comes at the cost of barely having privacy.

% What are the main approaches of previous works to achieve APPI? How are they
% lacking? If possible, use titled paragraphs to emphasize the main points.

From a cryptographic point of view, and combining both privacy and
identity/authentication requirements, group signatures (GS) and anonymous
credentials (AC) are surely among the most well known advanced schemes that
aim at addressing subsets of the previous problems. However, they typically fall
short of enabling utility or accountability without sacrificing privacy.

\paragraph{Limitations of Group Signatures.} %
For instance, in a vanilla GS scheme, a group of users can create a signature on
behalf of the group, without revealing the specific user who created it. At any
later point in time, an authority can extract from such group signature the
``identity'' of it signer (process called \emph{opening}). We find here two main
drawbacks:

\begin{enumerate}
\item This usually results in a sort of all-or-nothing approach. 
\item The concept of ``identity'' is barely defined.
\end{enumerate}

Concerning (1), either the signer is anonymous (before opening); or fully
deanonymized (after opening). Some variants on the traditional opening do not
fully deanonymize, but incorporate some other strict tradeoff. Like giving the
users full control \cite{dl21}, only allowing linking \needcite, or some
arbitrary extension such as allowing deanonymization only to the eyes of other
group members \needcite, or sequential linkability \cite{dl21}. While these
variants may be very suitable for concrete settings, it is not hard to see that
adapting them to other scenarios will be hard -- if possible at all.

Concerning (2), it is even surprising that, during the almost 30 years of
research in group signatures, the notion of ``identity'' of a group member
mostly reduces to an integer representing the order in which a user joined
the group -- or, at the most, some long-term public key. Thus, even with
notorious efforts to generalize accountability in group signatures,
re-identifying the signer from an index still forces to assume an external
ecosystem that endows that index with actual real world identyfing information.
Needless to say, this introduces further points of centralization, and
trust assumptions.

\paragraph{Limitations of Anonymous Credentials.} %
In a vanilla AC scheme, an (set of) issuer(s) grant credentials to users,
attesting for attributes that these users own. For intance, their age,
nationality, blood type, etc. Users can later prove arbitrary claims over
these attributes -- the most simple being selective disclosure; i.e., just
plainly revealing a subset of the attributes in a credential. Even from basic
selective disclosure, realistic applications can be built. The rationale being
that these attributes actually establish direct links with real identifying
information. However, accountability is barely considered, being blacklisting
of credentials almost the only explored variant.

While this may initially seem ok, closer thought quickly shows limitations. For
instance, think of a user who authenticates with an anonymous credential to get
in a night club. If the user proves being over the age limit, s/he will be let
in. However, if some time later, s/he is diagnosed positive in some contagious
disease, blacklisting does not serve any purpose. Incentivation mechanisms also
fit very well: a scenario in which a user authenticates anonymously to perform
some action that may receive some reward later will also not be implementable
with a credential scheme that only provides blacklisting. The problem is not
only practical, though. Without having suitable models to cater for these needs,
existing schemes just cannot be proven secure easily if they were to be used
to approximate solutions to the previous problems.

Given the previous, there is one first logical question: \emph{Can we come up
  with a scheme that combines the utility of AC with the accountability of
  GS?} But even then, we may run into a too rigid scheme. Thus, the follow-up
question would be: \emph{Can we design this combined scheme, such that it covers
a wide range of real world scenarios?}

%% 2 pages max for the previous; end with introduction (1-2 sentences) to what
%% would be ideal.

% Contributions first a small paragraph with an overview. Then one titled
% paragraph per contribution. Linking when possible with previous limitations.
\subsection{Our Contributions}
\label{ssec:contributions}

In this work, we make our attempt on addressing both questions.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "uas"
%%% End:

### 2022-05-10 prism-backend/research/gsac/uas.tex
Working on intro.

\begin{abstract}
  Much work has been done in the last 4 decades on privacy-preserving
  identities and authentication that tries to incorporate some sort of utility
  and accountability without losing (too much) privacy. Among the most relevant
  efforts we can find group signatures and anonymous credentials.
  %
  Anonymous credentials (AC) let users obtain attestations by trusted
  issuers on certain attributes -- thus implicitly defining the concept of
  identity as ``a set of attributes'' -- and allow credential owners to prove
  claims on these attributes. Indeed, this has proven to be very
  flexible, to the point that it has served as inspiration to design W3C's
  Verifiable Credentials, a relevant industry-led effort. However, the
  accountability offered by anonymous credentials is quite limited; frequently
  restricted to blacklisting, if present at all.
  %
  On the other hand, group signatures (GS) have traditionally focused on plain
  authentication without attributes (except very few exceptions), making instead
  more emphasis on the accountability side. That is, not only allowing full
  deanonymization, but also variants such as sequential linking, user-controlled
  anonymity, or even full unlinkability.
  %
  Not surprisingly, GS and AC share very common syntax, models and frequently
  also constructions.

  We argue that, for the real world, both functionalities are essential. Taking
  advantage of their similarities, we first propose a trivial combination, that
  we refer to with the original \GSAC acronym. Through \GSAC, we
  showcase the benefits of even a simple combination. We give a common syntax,
  security model, and generic construction for \GSAC.
  %
  Next, we acknowledge that even \GSAC has strong limitations. Namely, it is too
  strict in the rules one can apply during issuance of credentials, and signing
  and opening of the produced signatures. While generalizations along these
  lines have been explored independently in the GS and AC domains, to the best
  of our knowledge, no single scheme offers full flexibility.
  We define a syntax, security model, and give a generic
  construction for a scheme meeting these requirements. Given its high
  flexibility to cater for different needs, we name it Universal Anonymous
  Signatures (\UAS). Indeed, an \UAS scheme lets issuers decide on issuance
  policies, which easily enable advanced use cases such as delegation, or
  issuance conditioned identity-related claims. It can also be used to build
  schemes that allow signers prove arbitrary claims on their identities, as well
  as deciding on which should be the opening authority that can apply some
  commonly agreed deanonymization function over the signer's identity (which is
  now well-defined, given the incorporation of attributes from the AC domain).
  %
  The techniques and building blocks we leverage are well known. The challenge
  is, rather, on the definitional side.
\end{abstract}

>>>>

\begin{abstract}
  Much work has been done in the last 4 decades on privacy-preserving
  identities and authentication that tries to incorporate some sort of utility
  and accountability without losing (too much) privacy. Among the most relevant
  efforts we can find group signatures (GS) and anonymous credentials (AC).
  %
  AC let users obtain attestations by trusted
  issuers on certain attributes -- thus implicitly defining the concept of
  identity as ``a set of attributes'' -- and allow credential owners to prove
  claims on these attributes. Indeed, this has proven to be very
  flexible, to the point that it inspired the design of W3C's
  Verifiable Credentials, a relevant industry-led effort. However, the
  accountability in AC schemes is quite limited; frequently restricted to
  blacklisting, if present at all.
  %
  On the other hand, GS have traditionally focused on plain
  authentication without attributes (except very few exceptions), making instead
  more emphasis on the accountability side. That is, not only allowing full
  deanonymization, but also variants such as sequential linking, user-controlled
  anonymity, or even full unlinkability.
  %
  Not surprisingly, GS and AC share very common syntax, models and frequently
  also constructions.

  We argue that, in the real world, both functionalities are essential. Taking
  advantage of their similarities, we first propose a trivial combination, that
  we refer to with the imaginative \GSAC acronym. Through \GSAC, we showcase
  the benefits that even a simple combination brings. We give a common syntax,
  security model, and generic construction for \GSAC.
  %
  Next, we acknowledge that \GSAC has limitations. Namely, it is too
  strict in the rules one can apply during issuance of credentials, and signing
  and opening of the produced signatures. While generalizations along these
  lines have been explored independently in the GS and AC domains, to the best
  of our knowledge, no single scheme offers full flexibility.
  We define a syntax, security model, and give a generic
  construction for a scheme meeting these requirements. Given its high
  flexibility to cater for different needs, we name it Universal Anonymous
  Signatures (\UAS). Indeed, an \UAS scheme lets issuers decide on issuance
  policies, which easily enable advanced use cases such as delegation, or
  issuance conditioned identity-related claims. It can also be used to build
  schemes that allow signers prove arbitrary claims on their identities, as well
  as deciding on which should be the opening authority that can apply some
  commonly agreed deanonymization function over the signer's identity (which is
  now well-defined, given the incorporation of attributes from the AC domain).
  %
  The techniques and building blocks we leverage are well known. The challenge
  is, rather, on the definitional side. To further evidence the universality of
  \UAS, we show how more restricted schemes can be built from vanilla \UAS.
\end{abstract}
